---
title: "Known Marker Plot in Clusters"
author: "caoqi"
date: 
output: html_document
---


```{r include=FALSE}
# import library 
library(Seurat)
library(SingleCellExperiment)

knitr::opts_chunk$set(fig.width=20, fig.height=12) 

# color
library(RColorBrewer)
colors <- subset(brewer.pal.info,category=='qual')
colorqual <- c()
for(i in nrow(colors):1) colorqual <- c(colorqual, brewer.pal(colors[i,'maxcolors'], name = rownames(colors)[i]) )
colorqual <- c(colorqual,colorqual,colorqual)


uniqcol <- c('Blues','Greens','Greys','Oranges','Purples','Reds')
colors <- brewer.pal.info[uniqcol,]
colorseq <- c()
for(i in 1:nrow(colors)) colorseq <- c(colorseq, brewer.pal(colors[i,'maxcolors'], name = rownames(colors)[i]) )
colorseq <- colorseq[-2]
colorseq <- c(colorseq,colorseq,colorseq)

```



```{r include=FALSE}
# test data
# input_file <- 'Seurat_tmp.counts.scanpy.hvg2000_PC15_res1.h5ad.rds' # scanpy.sce.rds
# input_file <- '../Seurat_tmp.counts.seurat.hvg2000_PC15_res1.rds' # seurat.rds
# input_file <- 'Seurat_A113.counts.scanpy.hvg2000_PC15_res1.0.h5ad.rds'
# input_file <- 'Seurat_A113.counts.scanpy.hvg2000_PC20_res1.0.h5ad.rds'

# input file ------------
args = commandArgs(T)
input_format <- args[1]
input_file <- args[2]


if(input_format == 'seurat'){
  ## a. seurat 
seu_data <- readRDS( input_file )
seu_data
# seu_data$clusters <- seu_data$seurat_clusters
seu_data@reductions$UMAP <- Reductions(seu_data, 'umap')

}


if(input_format == 'sce'){
  ## b. scanpy.sce 
sce_data <- readRDS( input_file )
sce_data
assayNames(sce_data) <- 'counts'

# sce to seurat 
seu_data <- as.Seurat(sce_data, data = NULL)
seu_data$clusters <- seu_data$louvain

}



Idents(seu_data) <- seu_data$clusters
seu_data_gene <- rownames(seu_data)
seu_data@meta.data[,'Type'] <- gsub('A..._([A-Z_]*)[0-9]?$','\\1',seu_data$Sample)


## outfile -------
# output_file <- paste0(input_file,'.cluster_marker.html')

```


seurat data 
```{r}
input_file
seu_data
```

umap plot colored by clusters and samples
```{r fig.height = 8, fig.width = 16}
p1 <- DimPlot(seu_data, group.by = 'clusters', label = T, cols = colorqual) + NoLegend()
p2 <- DimPlot(seu_data, group.by = 'Sample', cols = colorqual, pt.size = 0.1) + NoLegend()
p1+p2
```


```{r fig.height = 6, fig.width = 16}
DimPlot(seu_data, group.by = 'Sample', cols = colorqual, pt.size = 0.1)
```



```{r fig.height = 6, fig.width = 7}
DimPlot(seu_data, group.by = 'Type', cols = colorqual)
```



UMAP plot showing the marker genes we used to distinguish major cell types:   
T: CD3D, CD3G;   
NK: NKG7;   
B: CD79A;   
dendritic cells: LILRA4, CLEC9A, CD1C, LAMP3;   
macrophages and monocytes: CD68, CD163;   
mast cells: KIT, CPA3;   
neutrophils: CSF3R, S100A8, S100A9;   
ILC: IL7R, RORA;   
fibroblast: COL1A1;   
Endothelial cells: VWF, CLEC4G;   
Parenchymal cells


# lymphocyte
### T and NK cell
```{r fig.height=12, fig.width=20}
features <- c('PTPRC','CD3D','CD3E','CD3G','CD8A','NKG7')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)

# MAIT
features <- c('CD4','SLC4A10','RORA','RORC','CXCR4')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```



### B cell 
```{r fig.height=12, fig.width=20}
# B cell: MS4A1, CD37
# plasma: CD79A, CD38, MZB1, SSR4
features <- c('MS4A1','CD37','CD79A','CD38','MZB1')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```



# Myeloid cell 
### neutrophil, Mast
```{r fig.height=12, fig.width=20}
features <- c('CSF3R','S100A8','S100A9','TPSAB1','CPA3','KIT')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)

features <- c('CSF1R','CSF2R','CSF3R','CXCL8','CXCL2','KIT')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)

```


### macrophage
```{r fig.height=12, fig.width=20}
features <- c('CD68','CD14','FCGR3A','MARCO','C1QC','SPP1')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```


### monocyte 
```{r fig.height=12, fig.width=20}
features <- c('FCN1','CD163','LILRA4', 'CLEC9A', 'CD1C', 'LAMP3')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```


# stromal cell 
### endothelial and Fibroblast cell
Usually, this cluster is very right.
```{r fig.height=12, fig.width=20}
features <- c('MGP', 'VWF', 'PLVAP','CLEC4G','COL1A1','ACTA2'  )
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```




# parenchymal 
### hepatocyte
```{r fig.height=12, fig.width=20}
features <- c("GPC3","TTR","AFP","ALB",'APOE','APOA1')
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```


### cholangio cell
```{r fig.height=12, fig.width=20}
features <- c("EPCAM","KRT7","PROM1","SOX9","KRT19")
FeaturePlot(seu_data, features = features, reduction = 'UMAP', ncol = 3, label = T)
```
